Model and results for identity signaling project

### 2020-05-23 paper/ms.tex
outline and some sections drafted for Model section

\vspace{1em}
\begin{table}[H]
  \centering
  \begin{tabular}{cl}
    Variable/parameter & Definition \\
    \toprule 
    $a_{ij}$   & Attitude of agent $i$ towards agent $j$; $a_i$ is the $N$-dimensional 
      attitude vector of agent $i$ \\
    $\tau_{ij}$ & Trait $j$ of agent $i$; $\tau_i$ is the $K$-dimensional trait vector of agent $i$ \\
    $K$      & Number of relevant traits determining similarity, liking, disliking, etc. \\
    $w$      & Homophily \\
    $d$      & Interaction penalty when one agent dislikes the other \\
    $\delta$ & Additional interaction penalty when both agents dislike each other \\
    $s$      & Added interaction benefit when agents are similar, as defined by their traits \\
    $S$      & Number of traits two agents must have in common to be considered similar \\
    $M$      & Number of traits used (if any) to define majority/minority populations \\
  \end{tabular}
  \caption{Model parameters.}
  \label{tab:params}
\end{table}

@@

\begin{table}[H]
  \begin{tabular}{ccp{5in}}
    Agent property & 
  \end{tabular}
\end{table}

Model individuals are agents who 
\begin{enumerate}
  \item have static ``in-born'' traits that determine pairwise similarity or dissimilarity
  \item hold attitudes about all other agents in the system
  \item have adopted a signaling and receiving strategy
\end{enumerate}Signaling strategies
may be either \emph{covert} or \emph{overt}, and receiving strategies may be
either \emph{generous} or \emph{churlish}. Unless otherwise noted, agents are
randomly initialized to be covert or overt with equal probability. The 
proportion of agents with specific signaling or receiving strategies will be
set to different values in the invasion experiments, explained in more detail below.

Each trait is represented arbitrarily by $\tau_{ij} \in \{-1, 1\}$ 
(MAYBE USE 'A' and 'B' INSTEAD SINCE THEY REALLY ARE CATEGORICAL?). 
Attitudes are $a_{ij} \in \{-1, 0, 1\}$ with meanings of 
disliking, neutral, or liking.

@@

Each model iteration consist of three stages. First, agents signal their identities
to some fraction of all other agents. It is at this step that agent attitudes
$a_{ij}$ are set. Whether or not an agent receives a signal depends on the
efficiency of overt and covert signaling, $E_{ov}$ and $E_{cov}$, respectively.
\emph{Efficiency} is the proportion of agents in the population that receive
a given agent's signal depending on whether the agent is an overt or covert
signaler.  Signals are sent and received once every model iteration. 
If and agent does receive a signal

@@

\begin{equation}
  \Pr(\text{$i$ interacts with $j$}) = \frac{1 + a_{ij} + a_{ji}}{2}
\end{equation}
\noindent

@@

\begin{table}[h]
  \centering
  \begin{tabular}{cccc}
    $a_{ij}$ & $a_{ji}$ & Similar? & Payoff $\pi_i = \pi_j$ \\
    \toprule
    1 & 1 & Yes & $1 + s$ \\
    1 & 0 & Yes & $1 + s$ \\
    1 & -1 & Yes & $1 + s - d$ \\
    1 & -1 & No & $1 + s - d$ \\
    0 & 0 & Yes & $1 + s$ \\
    0 & 0 & No & $1$ \\
    -1 & 0 & Yes & $1 + s - d$ \\
    -1 & 0 & No  & $1 - d$ \\
    -1 & -1 & Yes & $1 + s - 2d$ \\
    -1 & -1 & No  & $1 - 2d$ \\
  \end{tabular}
  \caption{Interaction payoff matrix. Certain conditions are impossible, and so are not
  shown. For instance, it is necessary for two agents to be similar for 
  one agent to like another agent (i.e.\ have $a_{ij} = 1$). Note that
  due to churlish receiving, two agents may be similar, but one dislikes the
  other while the other likes the one. There could be agents who dislike one
  another due to churlish receiving if they have never signaled each other.}
\end{table}

@@

All agents select another agent in the population to be a teacher. The 
learner agent probabilistically adopts its teacher's signaling or receiving 
strategy---which type of strategy might be learned is determined by a coin flip. 
The probability that the learner adopts the teacher's strategy is set 
calculated by the logistic sigmoid function applied to the quotient of 
teacher and learner payoffs accumulated in one signaling/receiving-interaction/payoff round.
Let $\pi_{i}$ be the accumulated payoff of agent $i$, and assume $i$ learns
from $j$. Set $p_{ij} = \pi_{j} / \pi_i$. 
The probability the learner will adopt the teacher's strategy is

>>>>

\vspace{1em}
\begin{table}[H]
  \centering
  \begin{tabular}{cl}
    Global parameter & Description \\
    \toprule 
    $K$      & Number of relevant traits determining similarity, liking, disliking, etc. \\
    $w$      & Homophily \\
    $d$      & Interaction penalty when one agent dislikes the other \\
    $\delta$ & Additional interaction penalty when both agents dislike each other \\
    $s$      & Added interaction benefit when agents are similar, as defined by their traits \\
    $S$      & Number of traits two agents must have in common to be considered similar \\
    $M$      & Number of traits used (if any) to define majority/minority populations \\
  \end{tabular}
  \caption{Global model parameters.}
  \label{tab:params}
\end{table}

@@

Model individuals are \emph{agents} who 
\begin{enumerate}
  \item have static ``in-born'' traits that determine pairwise similarity or dissimilarity
  \item hold attitudes about all other agents in the system
  \item have adopted a signaling and receiving strategy
\end{enumerate}Signaling strategies
may be either \emph{covert} or \emph{overt}, and receiving strategies may be
either \emph{generous} or \emph{churlish}. Unless otherwise noted, agents are
randomly initialized to be covert or overt with equal probability. The 
proportion of agents with specific signaling or receiving strategies will be
set to different values in the invasion experiments, explained in more detail below.

Agents have $K$ traits, where $K$ is a parameter representing the cultural 
complexity of the system. In the real world, different systems may have 
different cultural complexities. For instance, in United States politics, there
is increasingly little subtlety and increasing partisanship, so $K \approx1 $.
However, if we were trying to understand scholarly positions on some research
question, we may have $K$ be very large, or $K$ could be large if we consider
agents to be, e.g., nations instead of people (as in Axelrod 1997).

\begin{table}[H]
  \centering
  \begin{tabular}{cp{4.5in}}
    Agent property & Description \\
    \toprule
    $a_{ij}$   & Attitude of agent $i$ towards agent $j$; $a_i$ is the $N$-dimensional 
      attitude vector of agent $i$. \\
    $\tau_{ik}$ & Trait $k$ of agent $i$; $\tau_i$ is the $K$-dimensional trait vector of agent $i$. \\
    Signaling strategy & Either \emph{overt} or \emph{covert}. Overt signaling always
                         results in the communication of the signaler's traits,
                         while covert signalers only reveal their traits to 
                         ``similar'' others; dissimilar others do not receive a covert signal. \\
    Receiving strategy & Either \emph{churlish} or \emph{generous}. Churlish
                         receivers default to dislike other agents from whom they have
                         not received a signal. Generous receivers default to
                         a neutral attitude towards others from whom they have
                         not received a signal.
  \end{tabular}
  \caption{Agent attributes.}
  \label{tab:agentAttributes}
\end{table}

\subsubsection{Determining inter-agent similarity}

Each trait $j$ of agent $i$ takes a value $\tau_{ik} \in \{-1, 1\}$. $-1$ and 1
are used only to emphasize that there are two opposing traits.
Attitudes are $a_{ij} \in \{-1, 0, 1\}$ with meanings of 
disliking, neutral, or liking. Attitudes are set at the signaling and receiving
rounds, explained in more detail below. Briefly, if agent $i$ recieves a
trait signal from agent $j$, $a_{ij}$ will be 1 if the two agents have
similar traits, and $-1$ if $i$ and $j$ are sufficiently different. Similarity
is $s_{ij} = 1 - d(\tau_i, \tau_j)$ where $d(\tau_i, \tau_j)$ is the
Hamming distance between trait vectors. The Hamming distance is the fraction
of vector elements that are different. So, if $\tau_i = (1,~-1,~1)$ and
$\tau_j = (1,~-1,~-1)$, then $d(\tau_i, \tau_j) = 1/3$ and $s_{ij}=2/3$.
Two agents are ``similar'' if their similarity $s_{ij} \geq S$, where
$S$ is the similarity threshold parameter. Otherwise they are considered
dissimilar.

If two agents are similar, this enables agents to like one another. On the
other hand, dissimilar agents may come to dislike one another if they signal
their traits to one another. As explained further in the following section on
model dynamics, inter-agent attitudes determine the probability agents will
interact; the payoff received when a particular dyad interacts; and the
probability that one agent chooses another as a teacher.

@@

Each model iteration consist of three stages. First, agents signal their identities
to some fraction of all other agents. It is at this step that agent attitudes
$a_{ij}$ are set. Whether or not an agent receives a signal depends on the
efficiency of overt and covert signaling, $E_{ov}$ and $E_{cov}$, respectively.
\emph{Efficiency} is the proportion of agents in the population that receive
a given agent's signal depending on whether the agent is an overt or covert
signaler.  Signals are sent and received once every model iteration. 
If agent $i$ does receive a signal that agent updates its attitudes about
signaler $j$, then $i$ updates its attitudes about $j$ ($a_{ij}$), according to
the process outlined above.

@@

\begin{equation}
  \Pr(\text{$i$ interacts with $j$}) = \frac{1 + w(a_{ij} + a_{ji})}{2}
\end{equation}
\noindent

@@

\begin{table}[h]
  \centering
  \begin{tabular}{cccc}
    Similar? & $a_{ij}$ & $a_{ji}$ & Payoff $\pi_i = \pi_j$ \\
    \toprule
    Yes   & 1 & 1   & $1 + s$ \\
    Yes   & 1 & 0   & $1 + s$ \\
    Yes   & 1 & -1  & $1 + s - d$ \\
    Yes   & 0 & 0   & $1 + s$ \\
    Yes   & -1 & 0  & $1 + s - d$ \\
    Yes   & -1 & -1 & $1 + s - 2d$ \\
    \midrule
    No    & 1 & -1  & $1 + s - d$ \\
    No    & 0 & 0   & $1$ \\
    No    & -1 & 0  & $1 - d$ \\
    No    & -1 & -1 & $1 - 2d$ \\
  \end{tabular}
  \caption{Interaction payoff matrix. Certain conditions are impossible, and so are not
  shown. For instance, it is necessary for two agents to be similar for 
  one agent to like another agent (i.e.\ have $a_{ij} = 1$). Note that
  due to churlish receiving, two agents may be similar, but one dislikes the
  other while the other likes the one. There could be agents who dislike one
  another due to churlish receiving if they have never signaled each other.}
\end{table}

@@

All agents select another agent in the population to be a teacher.
Teacher selection is random, where the probability an agent selects another
agent is proportional to the probability of interaction given in 
Equation~\ref{eq:interactionProb}, normalized by the sum of all interaction
probabilities,

\begin{equation}
  \Pr(\text{$i$ chooses teacher $j$}) = 
    \frac{1 + w(a_{ij} + a_{ji})}{\sum_{j \neq i}(1 + w(a_{ij} + a_{ji}))}
\end{equation}
% \noindent

The 
learner agent probabilistically adopts its teacher's signaling or receiving 
strategy---which type of strategy might be learned is determined by a coin flip. 
The probability that the learner adopts the teacher's strategy is set 
calculated by the logistic sigmoid function applied to the quotient of 
teacher and learner payoffs accumulated in one
signaling/receiving-interaction/payoff round.
Let $\pi_{i}$ be the accumulated payoff of agent $i$, and assume $i$ learns
from $j$. Set $p_{ij} = \pi_{j} / \pi_i$. 
The probability the learner will adopt the teacher's strategy is
